Tracking and predicting turn structure during language acquisition

### 2014-09-10 paper/CasillasFrank-Anticip.tex
mcf changes

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/FIG-FL-stim.png}
\end{center}
\caption{Example frame from a conversation video used in Experiment 1.} 
\label{speakers}
\end{figure}

@@

\textit{Video-recordings}. We recorded pairs of talkers while they conversed in a sound-attenuated booth. Each talker was a native speaker of the language being recorded, and each talker pair was male-female. Using a Marantz PMD 660 solid state field recorder, we captured audio from two lapel microphones, one attached to each participant, while simultaneously recording video from the built-in camera of a MacBook. The talkers were volunteers, all members of the local post-graduate community, and were acquainted with their recording partner ahead of time. Each recording session began with a 20-minute warm-up period of spontaneous conversation during which the pair talked for five minutes on four topics (favorite foods, entertainment, hometown layout, and pets). Then we asked talkers to choose a new topic---one relevant to young children (e.g., riding a bike, eating breakfast)---and to improvise a dialogue on that topic. We asked them to speak as if they were on a children's television show in order to elicit child-directed speech toward each other. We recorded until the talkers achieved at least 30 seconds of uninterrupted discourse with enthusiastic, child-directed speech. Most talker pairs took less than five minutes to complete the task, usually by agreeing on a rough script at the start. We encouraged talkers to ask at least a few questions to each other during the improvisation. The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke a third or more language. We chose male-female pairs as a natural way of creating contrast between the two talker voices. See the videos here: NOTE: add link.}

@@

Because the conversational stimuli were recorded semi-spontaneously, the duration turn transitions and the number of speaker transitions in each video was variable. We measured the duration of each turn transition from the audio recording associated with each video. We excluded turn transitions longer than 550 msec and shorter than 0 msec (those with transitional overlap) from analysis.\footnote{Overlap occurs when an addressee begins a new turn before the current turn is finished. When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response begins \textit{earlier} than expected; participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}. As such, they would still be fixated on the prior speaker when the overlap started, and then would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$22) and non-English (N$=$21) videos. On average, the inter-turn gaps for English videos (mean$=$273 msec, median$=$291 msec) were slightly longer than for non-English videos (mean$=$226 msec, median$=$205 msec). The longer gaps in the English videos could give them a slight advantage: our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure 2), so participants had slightly more time to make anticipatory shifts in the English videos.

@@

\subsubsection{Procedure} 
Participants sat in front of an SMI 120Hz corneal reflection eye-tracker mounted beneath a large flatscreen display. The display and eye-tracker were secured to a table with an ergonomic arm that allowed the experimenter to position the whole apparatus at a comfortable height, approximately 60 cm from the viewer. We placed stereo speakers on the table, to the left and right of the display. Before the experiment started, we warned adult participants that they would see videos in several languages and that, though they weren't expected to understand the content of non-English videos, we \textit{would} ask them to answer general, non-language-based questions about the conversations. Then after each video we asked participants one of the following randomly-assigned questions: ``which speaker talked more?'', ``which speaker asked the most questions?'' ``which speaker seemed more friendly?'', and ``did the speakers' level of enthusiasm shift during the conversation?'' We also asked if the participants could understand any of what was said after each video. The participants responded verbally while an experimenter noted their responses.

Children were less inclined to simply sit and watch videos of conversation in languages they didn't speak, so we used a different procedure to keep them engaged. The experimenter started each session by asking the child about what languages he or she could speak, and about what other languages he or she had heard of. Then the experimenter expressed her own enthusiasm for learning about new languages, and invited the child to watch a video about ``new and different languages'' together. If the child agreed to watch, the experimenter and the child sat together in front of the display, with the child centered in front of the tracker and the experimenter off to the side. If the child began to look bored during video playback, the experimenter would talk during the filler videos, either commenting on the previous conversation (``That was a neat language!'') or giving the language name for the next conversation (``This next one is called Hebrew. Let's see what it's like.'') The experimenter's comments reinforced the video-watching as a joint task.

@@

\subsubsection{Data preparation and coding}
To determine whether participants predicted upcoming turn transitions, we needed to define a set of criteria for what counted as an anticipatory gaze shift. Prior work using similar experimental procedures has found that adults and children make anticipatory gaze shifts to upcoming talkers within a wide time frame; the earliest shifts occur before the end of the prior turn, and the latest occur after the onset of the response turn, with most shifts occurring in the inter-turn gap (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}). Following prior work, we measured how often our participants shifted their gaze from the prior to the upcoming speaker \textit{before} the shift in gaze could have been initiated in reaction to the onset of the speaker's response. In doing so, we assumed that it takes children 333 msec to plan an eye movement, following Fernald and colleagues' (2008) similar `looking while listening' method.

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.7\textwidth]{figures/FIG-AnticipCriteria.png}
\end{center}
\caption{Schematic summary of criteria for anticipatory gaze shifts from speaker A to speaker B during a turn transition. The criteria were the same for adults and children, except children were assumed to have a 333 msec window for planning eye movements, while adults were only given 200 msec.} 
\label{speakers}
\end{figure}

Using these criteria, we created a three-step algorithm for detecting anticipatory gaze switches. Consider, for example, when a child observes a turn transition between speaker A and speaker B (Figure 2). First, the child would need to fixate on A for at least 100 msec at the end of A's turn. But what is the relevant window for fixation? Consistent with the distribution of responses in explicit turn-boundary prediction tasks (de Ruiter et al., 2006), we counted any early switches as ``truly turn-related'' as long as they occurred in the final 333 msec of A's turn (200 msec for adults). To allow for these early switches, the window for fixating on A must start 433 msec before the end of A's turn (300 msec for adults). Second, given that the child fixated on A, we checked to see if the child also fixated on B. Because we allowed for early switches, the first possible fixation on B could start 333 msec before the end of A's turn. Assuming that it took the child 333 msec to plan an eye movement, the latest point when the gaze could have \textit{anticipated} speech would be 333 msec after B's turn began---a switch after this point could simply be a reaction to the onset of B's speech. Thus, the relevant window for fixating on B started 333 msec before the gap and ended 433 msec after it (200 msec and 300 ms, respectively, for adults). Success on the first two criteria jointly implied that the child made an anticipatory switch in gaze. Our third step was to estimate when the switch occurred. Starting at a point when we knew participants were fixated on A (from step 1), we incremented forward in time to the first gaze measurement where the child looked away from A. We recorded the look-away timestamp as the time of gaze launch to the upcoming speaker.

\textit{Random baseline analysis}. To ensure that participants' looking behavior was turn-related and not simply the result of random switching, we corrected for the probability of making an anticipatory switch at random. We estimated the baseline chance of making an anticipatory switch at each turn transition for each person. We ran all participants' eye-tracking data through switch identification with 100 randomly-shuffled versions of the original turn transition windows (Figure 3). For each of the 100 randomly-shuffled versions, we recorded and then averaged whether or not each participant made an anticipatory switch at each ``turn transition''. We then subtracted the random baseline value from each participant's actual switch result (1 or 0) for each turn transition to arrive at the final, baseline-corrected values. Subtracting the random baseline this way can give us an estimate of how often participants switched above chance across ages, conditions, and transition types. If participants make the same number of anticipatory switches, regardless of turn transition location, there would be no evidence that their switching behavior was linked to the conversation stimulus.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/FIG-ShuffledWindows.png}
\end{center}
\caption{Example of shuffling for five turn transition windows. The windows were +/-433 msec around the inter-turn gap for children, and from +/-300 msec for adults.} 
\label{shuffling}
\end{figure}

@@

Participants looked at the screen most of the time during video playback (81\% and 91\% on average for children and adults, respectively). They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: they gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table 1). Even three-year-olds looked more at the current speaker than anything else, whether the videos were in a language they could understand or not. Children looked at the current speaker less than adults did during the non-English videos. Despite this, their looks to the addressee did not increase substantially in the non-English videos, indicating that their looks away were probably related to boredom rather than confusion about ongoing turn structure. Overall, participants' pattern of gaze to current speakers indicated that they performed basic turn tracking during the videos, regardless of language.

\begin{table}[H]
\begin{center}
  \begin{tabular}{llcccc}
    \hline
    Age group & Condition & Speaker & Addressee & Other onscreen & Offscreen\\ 
    \hline
    3 & English & 0.61 & 0.16 & 0.14 & 0.08 \\ 
    4 & English & 0.60 & 0.15 & 0.11 & 0.13 \\ 
    5 & English & 0.57 & 0.15 & 0.16 & 0.12 \\ 
    Adult & English & 0.63 & 0.16 & 0.16 & 0.05 \\ 
    3 & Non-English & 0.38 & 0.17 & 0.20 & 0.25 \\ 
    4 & Non-English & 0.43 & 0.19 & 0.21 & 0.18 \\ 
    5 & Non-English & 0.40 & 0.16 & 0.26 & 0.18 \\ 
    Adult & Non-English & 0.58 & 0.20 & 0.16 & 0.07 \\ 
%   Overall & & 0.49 & 0.17 & 0.18 & 0.16 \\
    \hline
  \end{tabular}
\end{center}
  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk.}
\label{looktableexp1}
\end{table}

We identified anticipatory gaze switches for all 43 usable turn transitions, based on the criteria and algorithm outlined in Section 2.1.4. Because participants had greater access to lexical and syntactic information in English, we expected to see greater anticipation in the English videos than in the non-English videos. We also predicted that anticipation would be greater following questions compared to non-questions; questions have early cues to upcoming turn transition (e.g., \textit{wh-} words, subject-auxiliary inversion), and also make a next response immediately relevant. We also predicted that anticipatory looks would increase with development, along with children's increased linguistic competence.

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-randvsreal-FL.png}
\end{center}
\caption{Proportion anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages. Error bars represent the standard error of the mean.} 
\label{randvsrealFL}
\end{figure}

Participants at all ages made anticipatory gaze switches more often than would be expected by chance, whether they could understand the language in the video or not (Figure 4). Anticipatory switch behavior in the non-English videos could not have been based on lexicosyntactic information, so both adults and children must have relied on other sources of information to make their predictions (e.g., prosody and non-verbal signals).

Overall, a greater proportion of participants made anticipatory switches in the English videos (0.17) compared to the non-English videos (0.09). But the tendency to make more anticipatory switches for English than non-English videos disappeared with age. While anticipatory switches increased with age for non-English videos, they decreased with age for English videos, resulting in nearly equal looking behavior for the two language conditions for adults  (Figure 5). Otherwise, age-based differences in predictive gaze shifts were minor.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.5\textwidth]{figures/FIG-conditions-FL.png}
\end{center}
\caption{Proportion anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
\label{conditionsFL}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-QvsNQ-FL.png}
\end{center}
\caption{Proportion anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages. Error bars represent the standard error of the mean.} 
\label{questionsFL}
\end{figure}

Participants also showed an advantage for questions. Children made anticipatory switches twice as often when they heard a question, compared to a non-question (18.7\% and 7.3\%). But the advantage for questions was not universal---it depended on language condition and age. The advantage for questions didn't emerge until age five in the non-English videos (Figure 6).

We confirmed these findings in two linear mixed-effects models of baseline-corrected anticipatory switches \citep{lme4, R}. We built one model each for children and adults. The child model included condition (English vs. non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot compute reliable cross-language differences. Gaze behavior might be best for languages that have more structural overlap English. English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \cite{carlson2005}. We would need multiple items from each of the language to check for similarity effects of specific linguistic features.}, transition type (question vs. non-question), age, and duration of the inter-turn gap as predictors, with full interactions between condition, transition type, and age.  We included the duration of the inter-turn gap as a predictor since longer gaps also provide more opportunities to make anticipatory switches (Figure 2). We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}. The adult model was exactly the same as the children's, excluding age-related effects.

@@

Contrary to our expectation, there was no main effect of age (\textit{$\beta$}=-0.03, \textit{SE}=0.027, \textit{t}=-1.145). Children's anticipatory switches \textit{did} increase with age for the non-English videos, but it also decreased with age for the English videos, resulting in a significant interaction of age and language condition (\textit{$\beta$}=0.083, \textit{SE}=0.027, \textit{t}=3.063). We had expected that, as children got older, they would also show more anticipatory switches. But this is contradicted by decreased looks with age for English videos, which even extended to the adult data (Figure 5). One possible explanation is that older children and adults found the videos in English easy to comprehend, and so tracked the conversation less closely. Compared to three-year-olds, adults and older children spent more time looking away from the current speaker and addressee in the English conversations. In contrast, compared to three-year-olds, they spent more time looking at the current speaker and less time looking away in the non-English conversations (Table 1).

It is also possible that the timing of anticipatory gaze switches changes as children get older. That is, the moment at which observers chose to launch their gaze away from the current speaker might change with age. For example, children could have reacted more immediately to early cues in the turn (e.g., ``Which one...'') as they got older. If a change in timing were sufficiently large, we might not capture it in our analysis window, which was designed to capture switches close to the turn transition. To test for a change in switch timing, we would have to run new analyses on a larger set of turn transitions---ones with controlled early and late cues \citep[see, e.g.,][]{bogelsmagyariInPrep}. Within the current analysis windows there were some ``early switches'' (i.e., those launched before the end of the prior turn). But participants made approximately equal numbers of early switches, regardless of age and condition: 11--15\% (Figure 7). The lack of increased early switching with age might suggest that changes in timing are either larger than the analysis window, or are not at play.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-cumulative-FL.png}
\end{center}
\caption{Cumulative proportion of anticipatory gaze switches planned during the transition window, across conditions and ages. Age is indicated by hue (light to dark; darker = older). The solid line is the offset of the pre-gap turn, and the dashed line is the median onset of the post-gap turn in each condition. ``Early switches'' occur before 0.} 
\label{cumulativeFL}
\end{figure}

Children made more anticipatory switches after hearing questions than non-questions, but the advantage for questions was driven by age and linguistic access. There was no main effect of transition type alone (\textit{$\beta$}=-0.209, \textit{SE}=0.151, \textit{t}=-1.384). Rather, the model confirmed a two-way interaction of linguistic condition and transition type (\textit{$\beta$}=0.445, \textit{SE}=0.19, \textit{t}=2.344). Presumably, question effects were larger in the English videos than the non-English videos because there were more cues to questionhood in English (i.e., non-verbal, prosodic, \textit{and} lexicosyntactic). Children younger than five did not distinguish questions from non-questions in the non-English videos (Figure 6). The children's model confirmed a three-way interaction of age, linguistic condition, and transition type (\textit{$\beta$}=-0.091, \textit{SE}=0.042, \textit{t}=-2.183). Lexicosyntactic information might then be critical for young children to identify questions in conversation. Cues like subject-auxiliary inversion, and \textit{wh}-words occur early in the turn and are clear indicators that an answer will come next. Importantly, the turns marked as ``questions'' in the non-English data were not real questions, but turns that \textit{sounded like} questions to native English speakers. Question-like prosodic cues in the non-English videos could have been markedly different from question prosody in English, making it more difficult for three- and four-year-olds to identify ``questions'' in those conversations.

@@

A few results from Study 1 were unexpected. First, we found no significant main effect of age in the children's data, and we saw very little difference between adults and children (besides the age-related interactions). It appears that, especially in English, children were as good at tracking the current speaker and anticipating upcoming turns as adults were. To find out when children first begin making spontaneous anticipatory looks to responders, we will have to look at younger children. Second, children, but not adults, showed an advantage for lexicosyntactic information in their anticipatory switches. An advantage for lexicosyntactic information is consistent with prior work on adult turn taking \citep{de-ruiter2006}, so it's curious that the adults did not show a difference in anticipatory switching between videos they did and did not understand. As mentioned, they may have effectively used the non-verbal cues in the videos to track the interactions fluently and, consequently, found the interactions so easy to follow that they did not track the turn-taking closely. To identify which \textit{linguistic} cues adults and children rely on, we have to control the presence of non-verbal cues in the videos.

>>>>

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/FIG-FL-stim.png}
\end{center}
\caption{Example frame from a conversation video used in Experiment 1.} 
\label{fig:speakers}
\end{figure}

@@

\textit{Video recordings}. We recorded pairs of talkers while they conversed in a sound-attenuated booth (see sample frame in Figure \ref{fig:speakers}). Each talker was a native speaker of the language being recorded, and each talker pair was male-female. Using a Marantz PMD 660 solid state field recorder, we captured audio from two lapel microphones, one attached to each participant, while simultaneously recording video from the built-in camera of a MacBook. The talkers were volunteers, all members of the local post-graduate community, and were acquainted with their recording partner ahead of time.

Each recording session began with a 20-minute warm-up period of spontaneous conversation during which the pair talked for five minutes on four topics (favorite foods, entertainment, hometown layout, and pets). Then we asked talkers to choose a new topic---one relevant to young children (e.g., riding a bike, eating breakfast)---and to improvise a dialogue on that topic. We asked them to speak as if they were on a children's television show in order to elicit child-directed speech toward each other. We recorded until the talkers achieved at least 30 seconds of uninterrupted discourse with enthusiastic, child-directed speech. Most talker pairs took less than five minutes to complete the task, usually by agreeing on a rough script at the start. We encouraged talkers to ask at least a few questions to each other during the improvisation. The resulting conversations were therefore not entirely spontaneous, but were as close as possible while still remaining child-oriented in topic, prosodic pattern, and lexicosyntactic construction.\footnote{All of the non-English talkers were fluent in English as a second language, and some fluently spoke a third or more language. We chose male-female pairs as a natural way of creating contrast between the two talker voices. See the videos here: NOTE: add link.}

@@

Because the conversational stimuli were recorded semi-spontaneously, the duration turn transitions and the number of speaker transitions in each video was variable. We measured the duration of each turn transition from the audio recording associated with each video. We excluded turn transitions longer than 550 msec and shorter than 0 msec (those with transitional overlap) from analysis.\footnote{Overlap occurs when an addressee begins a new turn before the current turn is finished. When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response begins \textit{earlier} than expected; participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}. As such, they would still be fixated on the prior speaker when the overlap started, and then would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$22) and non-English (N$=$21) videos. On average, the inter-turn gaps for English videos (mean$=$273 msec, median$=$291 msec) were slightly longer than for non-English videos (mean$=$226 msec, median$=$205 msec). The longer gaps in the English videos could give them a slight advantage: our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.

@@

\subsubsection{Procedure} 
Participants sat in front of an SMI 120Hz corneal reflection eye-tracker mounted beneath a large flatscreen display. The display and eye-tracker were secured to a table with an ergonomic arm that allowed the experimenter to position the whole apparatus at a comfortable height, approximately 60 cm from the viewer. We placed stereo speakers on the table, to the left and right of the display.

Before the experiment started, we warned adult participants that they would see videos in several languages and that, though they weren't expected to understand the content of non-English videos, we \textit{would} ask them to answer general, non-language-based questions about the conversations. Then after each video we asked participants one of the following randomly-assigned questions: ``which speaker talked more?'', ``which speaker asked the most questions?'' ``which speaker seemed more friendly?'', and ``did the speakers' level of enthusiasm shift during the conversation?'' We also asked if the participants could understand any of what was said after each video. The participants responded verbally while an experimenter noted their responses.

Children were less inclined to simply sit and watch videos of conversation in languages they didn't speak, so we used a different procedure to keep them engaged. The experimenter started each session by asking the child about what languages he or she could speak, and about what other languages he or she had heard of. Then the experimenter expressed her own enthusiasm for learning about new languages, and invited the child to watch a video about ``new and different languages'' together. If the child agreed to watch, the experimenter and the child sat together in front of the display, with the child centered in front of the tracker and the experimenter off to the side. If the child began to look bored during video playback, the experimenter would talk during the filler videos, either commenting on the previous conversation (``That was a neat language!'') or giving the language name for the next conversation (``This next one is called Hebrew. Let's see what it's like.''). The experimenter's comments reinforced the video-watching as a joint task.

@@

\subsubsection{Data preparation and coding}
\label{sec:algorithm}

To determine whether participants predicted upcoming turn transitions, we needed to define a set of criteria for what counted as an anticipatory gaze shift. Prior work using similar experimental procedures has found that adults and children make anticipatory gaze shifts to upcoming talkers within a wide time frame; the earliest shifts occur before the end of the prior turn, and the latest occur after the onset of the response turn, with most shifts occurring in the inter-turn gap (Keitel et al., 2013; Hirvenkari, 2013; Tice and Henetz, \citeyear{TiceHenetz11}). Following prior work, we measured how often our participants shifted their gaze from the prior to the upcoming speaker \textit{before} the shift in gaze could have been initiated in reaction to the onset of the speaker's response. In doing so, we assumed that it takes children 333 msec to plan an eye movement, following Fernald and colleagues' (2008) similar `looking while listening' method.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.7\textwidth]{figures/FIG-AnticipCriteria.png}
\end{center}
\caption{Schematic summary of criteria for anticipatory gaze shifts from speaker A to speaker B during a turn transition. The criteria were the same for adults and children, except children were assumed to have a 333 msec window for planning eye movements, while adults were only given 200 msec.} 
\label{fig:criterion}
\end{figure}

Using these criteria, we created a three-step algorithm for detecting anticipatory gaze switches. Consider, for example, when a child observes a turn transition between speaker A and speaker B (Figure \ref{fig:criterion}). First, the child would need to fixate on A for at least 100 msec at the end of A's turn. But what is the relevant window for fixation? Consistent with the distribution of responses in explicit turn-boundary prediction tasks (de Ruiter et al., 2006), we counted any early switches as ``truly turn-related'' as long as they occurred in the final 333 msec of A's turn (200 msec for adults). To allow for these early switches, the window for fixating on A must start 433 msec before the end of A's turn (300 msec for adults). Second, given that the child fixated on A, we checked to see if the child also fixated on B. Because we allowed for early switches, the first possible fixation on B could start 333 msec before the end of A's turn. Assuming that it took the child 333 msec to plan an eye movement, the latest point when the gaze could have \textit{anticipated} speech would be 333 msec after B's turn began---a switch after this point could simply be a reaction to the onset of B's speech. Thus, the relevant window for fixating on B started 333 msec before the gap and ended 433 msec after it (200 msec and 300 ms, respectively, for adults). Success on the first two criteria jointly implied that the child made an anticipatory switch in gaze. Our third step was to estimate when the switch occurred. Starting at a point when we knew participants were fixated on A (from step 1), we incremented forward in time to the first gaze measurement where the child looked away from A. We recorded the look-away timestamp as the time of gaze launch to the upcoming speaker.

\paragraph{Random baseline analysis} To ensure that participants' looking behavior was turn-related and not simply the result of random switching, we corrected for the probability of making an anticipatory switch at random. We estimated the baseline chance of making an anticipatory switch at each turn transition for each person. We ran all participants' eye-tracking data through switch identification with 100 randomly-shuffled versions of the original turn transition windows (Figure \ref{fig:shuffling}). For each of the 100 randomly-shuffled versions, we recorded and then averaged whether or not each participant made an anticipatory switch at each ``turn transition.'' We then subtracted the random baseline value from each participant's actual switch result (1 or 0) for each turn transition to arrive at the final, baseline-corrected values. Subtracting the random baseline this way can give us an estimate of how often participants switched above chance across ages, conditions, and transition types. If participants made the same number of anticipatory switches, regardless of turn transition location, there would be no evidence that their switching behavior was linked to the conversation stimulus.

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/FIG-ShuffledWindows.png}
\end{center}
\caption{Example of shuffling for five turn transition windows. The windows were +/-433 msec around the inter-turn gap for children, and from +/-300 msec for adults.} 
\label{fig:shuffling}
\end{figure}

@@

\subsubsection{Descriptive analysis}

Participants looked at the screen most of the time during video playback (81\% and 91\% on average for children and adults, respectively). They primarily kept their eyes on the person who was currently speaking in both English and non-English videos: they gazed at the current speaker between 38\% and 63\% of the time, looking back at the addressee between 15\% and 20\% of the time (Table \ref{tab:e1_look}). Even three-year-olds looked more at the current speaker than anything else, whether the videos were in a language they could understand or not. Children looked at the current speaker less than adults did during the non-English videos. Despite this, their looks to the addressee did not increase substantially in the non-English videos, indicating that their looks away were probably related to boredom rather than confusion about ongoing turn structure. Overall, participants' pattern of gaze to current speakers indicated that they performed basic turn tracking during the videos, regardless of language.

\begin{table}[t]
\begin{center}
  \begin{tabular}{llcccc}
    \hline
    Age group & Condition & Speaker & Addressee & Other onscreen & Offscreen\\ 
    \hline
    3 & English & 0.61 & 0.16 & 0.14 & 0.08 \\ 
    4 & English & 0.60 & 0.15 & 0.11 & 0.13 \\ 
    5 & English & 0.57 & 0.15 & 0.16 & 0.12 \\ 
    Adult & English & 0.63 & 0.16 & 0.16 & 0.05 \\ 
    3 & Non-English & 0.38 & 0.17 & 0.20 & 0.25 \\ 
    4 & Non-English & 0.43 & 0.19 & 0.21 & 0.18 \\ 
    5 & Non-English & 0.40 & 0.16 & 0.26 & 0.18 \\ 
    Adult & Non-English & 0.58 & 0.20 & 0.16 & 0.07 \\ 
%   Overall & & 0.49 & 0.17 & 0.18 & 0.16 \\
    \hline
  \end{tabular}
\end{center}
  \caption{Average proportion of gaze to the current speaker and addressee during periods of talk.}
\label{tab:e1_look}
\end{table}

We identified anticipatory gaze switches for all 43 usable turn transitions, based on the criteria and algorithm outlined in Section \ref{sec:algorithm}. Because participants had greater access to lexical and syntactic information in English, we expected to see greater anticipation in the English videos than in the non-English videos. We also predicted that anticipation would be greater following questions compared to non-questions; questions have early cues to upcoming turn transition (e.g., \textit{wh-} words, subject-auxiliary inversion), and also make a next response immediately relevant. We also predicted that anticipatory looks would increase with development, along with children's increased linguistic competence.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-randvsreal-FL.png}
\end{center}
\caption{Proportion anticipatory gaze switches made for actual (solid) and randomly-shuffled (dashed) turn transitions in each condition, and across ages. Error bars represent the standard error of the mean.} 
\label{fig:randvsrealFL}
\end{figure}

Participants at all ages made anticipatory gaze switches more often than would be expected by chance, whether they could understand the language in the video or not (Figure \ref{fig:randvsrealFL}). Anticipatory switch behavior in the non-English videos could not have been based on lexicosyntactic information, so both adults and children must have relied on other sources of information to make their predictions (e.g., prosody and non-verbal signals).

Overall, a greater proportion of participants made anticipatory switches in the English videos (0.17) compared to the non-English videos (0.09). But the tendency to make more anticipatory switches for English than non-English videos disappeared with age. While anticipatory switches increased with age for non-English videos, they decreased with age for English videos, resulting in nearly equal looking behavior for the two language conditions for adults. Otherwise, age-based differences in predictive gaze shifts were minor (Figure \ref{fig:conditionsFL}).

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.5\textwidth]{figures/FIG-conditions-FL.png}
\end{center}
\caption{Baseline-corrected proportion anticipatory gaze switches made for turn transitions in each language condition, and across ages. Error bars represent the standard error of the mean.} 
\label{fig:conditionsFL}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-QvsNQ-FL.png}
\end{center}
\caption{Proportion anticipatory gaze switches made for question- (dashed) and non-question- (dotted) turn transitions in each condition, and across ages. Error bars represent the standard error of the mean.} 
\label{fig:questionsFL}
\end{figure}

Participants also showed an advantage for questions. Children made anticipatory switches twice as often when they heard a question, compared to a non-question (18.7\% and 7.3\%). But the advantage for questions was not universal---it depended on language condition and age. The advantage for questions didn't emerge until age five in the non-English videos (Figure \ref{fig:questionsFL}).

\subsubsection{Statistical analysis}

We quantified these findings using two linear mixed-effects models of baseline-corrected anticipatory switches \citep{lme4, R}. We built one model each for children and adults. The child model included condition (English vs. non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot compute reliable cross-language differences. Gaze behavior might be best for languages that have more structural overlap English. English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \cite{carlson2005}. We would need multiple items from each of the language to check for similarity effects of specific linguistic features.}, transition type (question vs. non-question), age, and duration of the inter-turn gap as predictors, with full interactions between condition, transition type, and age.  We included the duration of the inter-turn gap as a predictor since longer gaps also provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}). We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}. The adult model was exactly the same as the children's, excluding age-related effects.

@@

Contrary to our expectation, there was no main effect of age (\textit{$\beta$}=-0.03, \textit{SE}=0.027, \textit{t}=-1.145). Children's anticipatory switches \textit{did} increase with age for the non-English videos, but it also decreased with age for the English videos, resulting in a significant interaction of age and language condition (\textit{$\beta$}=0.083, \textit{SE}=0.027, \textit{t}=3.063). We had expected that, as children got older, they would also show more anticipatory switches. But this is contradicted by decreased looks with age for English videos, which even extended to the adult data. One possible explanation is that older children and adults found the videos in English easy to comprehend, and so tracked the conversation less closely. Compared to three-year-olds, adults and older children spent more time looking away from the current speaker and addressee in the English conversations. In contrast, compared to three-year-olds, they spent more time looking at the current speaker and less time looking away in the non-English conversations (Table \ref{tab:e1_look}).

It is also possible that the timing of anticipatory gaze switches changes as children get older. That is, the moment at which observers chose to launch their gaze away from the current speaker might change with age. For example, children could have reacted more immediately to early cues in the turn (e.g., ``Which one...'') as they got older. If a change in timing were sufficiently large, we might not capture it in our analysis window, which was designed to capture switches close to the turn transition. To test for a change in switch timing, we would have to run new analyses on a larger set of turn transitions---ones with controlled early and late cues \citep[see, e.g.,][]{bogelsmagyariInPrep}. Within the current analysis windows there were some ``early switches'' (i.e., those launched before the end of the prior turn). But participants made approximately equal numbers of early switches, regardless of age and condition: 11--15\% (Figure \ref{fig:cumulativeFL}). The lack of increased early switching with age might suggest that changes in timing are either larger than the analysis window, or are not at play.

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/FIG-cumulative-FL.png}
\end{center}
\caption{Cumulative proportion of anticipatory gaze switches planned during the transition window, across conditions and ages. Age is indicated by hue (light to dark; darker = older). The solid line is the offset of the pre-gap turn, and the dashed line is the median onset of the post-gap turn in each condition. ``Early switches'' occur before 0.} 
\label{fig:cumulativeFL}
\end{figure}

Children made more anticipatory switches after hearing questions than non-questions, but the advantage for questions was driven by age and linguistic access. There was no main effect of transition type alone (\textit{$\beta$}=-0.209, \textit{SE}=0.151, \textit{t}=-1.384). Rather, the model confirmed a two-way interaction of linguistic condition and transition type (\textit{$\beta$}=0.445, \textit{SE}=0.19, \textit{t}=2.344). Presumably, question effects were larger in the English videos than the non-English videos because there were more cues to questionhood in English (i.e., non-verbal, prosodic, \textit{and} lexicosyntactic). Children younger than five did not distinguish questions from non-questions in the non-English videos (Figure \ref{fig:questionsFL}). The children's model confirmed a three-way interaction of age, linguistic condition, and transition type (\textit{$\beta$}=-0.091, \textit{SE}=0.042, \textit{t}=-2.183). Lexicosyntactic information might then be critical for young children to identify questions in conversation. Cues like subject-auxiliary inversion, and \textit{wh}-words occur early in the turn and are clear indicators that an answer will come next. Importantly, the turns marked as ``questions'' in the non-English data were not real questions, but turns that \textit{sounded like} questions to native English speakers. Question-like prosodic cues in the non-English videos could have been markedly different from question prosody in English, making it more difficult for three- and four-year-olds to identify ``questions'' in those conversations.

@@

\subsubsection{Summary}

Several results from Experiment 1 were unexpected. First, we found no significant main effect of age in the children's data, and we saw very little difference between adults and children (besides the age-related interactions). It appears that, especially in English, children were as good at tracking the current speaker and anticipating upcoming turns as adults were. To find out when children first begin making spontaneous anticipatory looks to responders, younger participants are required. Experiment 2 thus includes a sample of one- and two-year-olds.

Second, children, but not adults, showed an advantage for lexicosyntactic information in their anticipatory switches. An advantage for lexicosyntactic information is consistent with prior work on adult turn taking \citep{de-ruiter2006}, so it's curious that the adults did not show a difference in anticipatory switching between videos they did and did not understand. As mentioned above, it is likely that they used non-verbal cues to track the interactions and consequently found the videos so easy to follow that they did not track the turn-taking closely. To identify which \textit{linguistic} cues adults and children rely on, it is necessary to control the presence of non-verbal cues in the stimulus---Experiment 2 implements this control.
